---
title: "SBP in Young Adults"
author: "Simon Bond"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up

First load in the data (I slightly edited it) and load some libraries

```{r cars}
library(tidyverse)
library(nlme)
library(rstan)
library(bayesplot)

df <- read.csv(file="Central SBP in young ISH_v2.csv") %>% 
  cctu::clean_names() %>% 
  mutate(se=sd/sqrt(n),
         weights=se^2
         )
summary(df)

```

# Basic analysis

This start with the assumption that $y_i \sim N( \mu, \sigma_{i}^2/n)$ in the hypertensive studies, 
and $y_i \sim N( \mu+ \delta, \sigma_{i}^2/n)$  in the optimal group. The standard deviation is $\sigma$,  and is assumed here to be a known fixed parameter. Also there is no heterogeniety between studies, akin to assuming you simply took seperate samples with identicial treatments,  from and identical population,  for each study. 

A basic frequentest analysis can be done that uses the fixed standard deviation. 


```{r gls, echo=FALSE}
m0 <- gls(mean~group, weights=~weights, data=df,control=glsControl(sigma=1)) 
m0 %>%  summary
```


# Extensions

We want to relax the assumptions made above. 

1. acknowledge there is sample variability in the standard deviations. IN fact theory says $n \hat{\sigma}^2 \sim \sigma_{within}^2 \Chi^2_{n-1}$
2. Allow the location to vary randomly between studies by $\eta_{i} \sim N(0, \sigma^2_{between})$
3. Allow also the between-patient SD to vary between studies  $\sigma_{within,i} \sim \Gamma(  \sigma_{within}\nu, \nu )$

These need Bayesian MCMC software to fit.  The R code is shown below, which in turn relies on individual files, contained in the folder  ( m0.stan, m1.stan, m2.stan, m4.stan).


```{r mcmc, eval=FALSE }


df_stan <- list(
  N = nrow(df),
  y=df$mean,
  x=1*(df$group=="optimal"),
  se=df$se,
  sample=df$n
)


fit0 <- stan(
  file = "m0.stan",  # Stan program
  data = df_stan,    # named list of data
  chains = 4,             # number of Markov chains
  warmup = 1000,          # number of warmup iterations per chain
  iter = 2000,            # total number of iterations per chain
  cores = 2,              # number of cores (could use one per chain)
  refresh = 100             # no progress shown
)
  
fit1 <- stan(
  file = "m1.stan",  # Stan program
  data = df_stan,    # named list of data
  chains = 4,             # number of Markov chains
  warmup = 1000,          # number of warmup iterations per chain
  iter = 2000,            # total number of iterations per chain
  cores = 2,              # number of cores (could use one per chain)
  refresh = 100             # no progress shown
)
save.image("fits.Rdata")
load("fits.Rdata")

fit2 <- stan(
  file = "m2.stan",  # Stan program
  data = df_stan,    # named list of data
  chains = 4,             # number of Markov chains
  warmup = 1000,          # number of warmup iterations per chain
  iter = 2000,            # total number of iterations per chain
  cores = 2,              # number of cores (could use one per chain)
  refresh = 100             # no progress shown
)

summary(fit2, pars=c("mu","delta","sigma_within","sigma_between"))

renv::install("bayesplot")
library(bayesplot)
mcmc_areas(fit3,pars=my_pars)
mcmc_areas(fit3,pars="sigma_within")
mcmc_areas(fit3,pars="delta")
traceplot(fit3, pars=my_pars)
summary(fit3, c(my_pars,"sigma_within_se"))
summary(fit2, pars=c("mu","delta","sigma_within","sigma_between"))



fit4 <- stan(
  file = "m4.stan",  # Stan program
  data = df_stan,    # named list of data
  chains = 4,             # number of Markov chains
  warmup = 2000,          # number of warmup iterations per chain
  iter = 10000,            # total number of iterations per chain
  cores = 2,              # number of cores (could use one per chain)
  refresh = 500             # no progress shown
)

save.image("fits.Rdata")
```

We ran the code earliear (each fit takes abotu 5 mins) saved the results, which is now re-loaded

In turn this 

1. Replicates the analysis above
2. Acknowledges sampling variability in the standard deviations
3. Adds in between-study heterogeneity in the location 
4.  Adds in between-study heterogeneity in the variance

```{r inspect}
load("fits.Rdata")


summary(fit0, pars=c("mu","delta"))$summary %>% knitr::kable(digits = 2)
summary(fit1, pars=c("mu","delta","sigma_within"))$summary%>% knitr::kable(digits = 2)
summary(fit2, pars=c("mu","delta","sigma_within", "sigma_between"))$summary%>% knitr::kable(digits = 2)
summary(fit4, pars=c("mu","delta","sigma_within","sigma_between","sigma_within_se"))$summary%>% knitr::kable(digits = 2)

```


# Comments

These all assume a flat prior for all the parameters. This is the lazy unthinking working assumption. 

There is evidence to support heterogeniety being present for both the location and variance. 

The difference between the groups is statistically significant (for the frequentist first model). The other models are Bayesian, and so p-values don't exist, but give credible intervals well away from zero.  They are all centred around 26-28 point estimates for the difference, the optimal group being lower.  The credible intervals widen, but nonethelss exclude 0,  for the models that incorporate the heterogeniety in location, which has a standard deviation of 9 ( `sigma_between` models 2/4),  comparable in size to the between-patient SD (`sigma_within` around 7.5). 



